---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mskcc.oncotree

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/mskcc.oncotree)](https://CRAN.R-project.org/package=mskcc.oncotree)
<!-- badges: end -->

The goal of `{mskcc.oncotree}` is to facilitate access to the [OncoTree
API](http://oncotree.mskcc.org/).

The OncoTree is an open-source ontology that was developed at Memorial Sloan
Kettering Cancer Center (MSK) for standardizing cancer type diagnosis from a
clinical perspective by assigning each diagnosis a unique OncoTree code.

Currently, the functionality provided is the retrieval of tumor types and the
mapping of tumor type codes across a few tumor type classification
systems.

## Installation

Install `{mskcc.oncotree}` from CRAN:

``` r
install.packages("mskcc.oncotree")
```

You can install the development version of `{mskcc.oncotree}` with:

``` r
# install.packages("remotes")
remotes::install_github("maialab/mskcc.oncotree")
```

## OncoTree tumor types

Get the tumor types defined by OncoTree in their latest release:

```{r tumor_types}
library(mskcc.oncotree)

(tumor_types <- get_tumor_types())
```

The mapping to the corresponding tumor type identifier(s) in the Uniﬁed Medical
Language System (UMLS) and in the National Cancer Institute (NCI) Thesaurus
classification systems is provided in columns `umls_code` and `nci_code`.

```{r mappings}
tumor_types[c('oncotree_code', 'umls_code', 'nci_code')]
```

`umls_code` and `nci_code` are list-columns because the mapping can be
one-to-many.

If you want, you may convert these columns to type `character` by unnesting with
`tidyr::unnest()`. This will have the side effect of creating more than one row
per tumor type for cases where the mapping is one-to-many.

```{r unnest}
tumor_types[c('oncotree_code', 'umls_code')] %>%
  tidyr::unnest(cols = 'umls_code', keep_empty = TRUE)
```

## OncoTree versions

The data provided by OncoTree is versioned. You may list the released versions
so far with `get_versions()`:

```{r versions}
get_versions()
```

The function `get_tumor_types()` accepts a release version, allowing to get data
for past versions of OncoTree:

```{r old_version}
get_tumor_types(oncotree_version = 'oncotree_2019_08_01')
```

## Mapping of tumor type codes

### UMLS and NCIt

OncoTree provides a few mappings between their identifiers (codes) and other
tumor classification systems. Currently, through their web API, only mappings
between OncoTree codes and Uniﬁed Medical Language System (UMLS) and National
Cancer Institute (NCI) Thesaurus (NCIt) codes are provided.

From `{mskcc.oncotree}` we provide four functions:

| Function             | Description                |
| -------------------- | -------------------------- |
| `oncotree_to_nci()`  | Map OncoTree to NCIt codes |
| `nci_to_oncotree()`  | Map NCIt to OncoTree codes |
| `oncotree_to_umls()` | Map OncoTree to UMLS codes |
| `umls_to_oncotree()` | Map UMLS to OncoTree codes |

### From and to NCIt

```{r api_mappings_nci}
oncotree_to_nci(c('MMB', 'PAOS'), expand = TRUE)

nci_to_oncotree(c('C3706', 'C8969'), expand = TRUE)
```

### From and to UMLS

```{r api_mappings_umls}
oncotree_to_umls(c('MMB', 'PAOS'), expand = TRUE)

umls_to_oncotree(c('C0205833', 'C0206642', 'C1334708'), expand = TRUE)
```

### Other mappings

Besides UMLS and NCIt, you may also map OncoTree codes to
[ICD-O](https://en.wikipedia.org/wiki/International_Classification_of_Diseases_for_Oncology)
topography and morphology, and [HemeOnc](https://hemonc.org/) codes. This
functionality is based on the file
[ontology_mappings.txt.](https://github.com/cBioPortal/oncotree/blob/master/scripts/ontology_to_ontology_mapping_tool/ontology_mappings.txt)
provided by OncoTree at their GitHub repository.

In OncoTree's discussion forum the developers have warned that this data can't
be expected to be neither accurate nor complete. However, it might still be
useful for some users, and hence we also provide access to these mappings with
the function `map_ontology_code()`.

Here are some usage examples `map_ontology_code()`:

```{R}
# Simple example
map_ontology_code(code = 'MMB', from = 'oncotree_code', to = 'nci_code')

# Omit the `code` argument to get all possible mappings. Note that
# one-to-many mappings will generate more than one row per `from` code.
map_ontology_code(from = 'oncotree_code', to = 'nci_code')

# Some mappings are one-to-many, e.g. "SRCCR", which means repeated rows for
# the same input code.
map_ontology_code(code = 'SRCCR', from = 'oncotree_code', to = 'nci_code')

# Using the `collapse` argument to "collapse" one-to-many mappings makes sure
# that the output has as many rows as the `from` vector.
map_ontology_code(code = 'SRCCR',
                  from = 'oncotree_code',
                  to = 'nci_code',
                  collapse = toString)

map_ontology_code(code = 'SRCCR',
                  from = 'oncotree_code',
                  to = 'nci_code',
                  collapse = list)

map_ontology_code(
  code = 'SRCCR',
  from = 'oncotree_code',
  to = 'nci_code',
  collapse = \(x) paste(x, collapse = ' ')
)

# `map_ontology_code()` is vectorized over `code`
map_ontology_code(code = c('AASTR', 'MDEP'), from = 'oncotree_code', to = 'nci_code')

# Map from ICDO topography to ICDO morphology codes
map_ontology_code(code = 'C72.9', from = 'icdo_topography_code', to = 'icdo_morphology_code')
```
